First, I thank my CMS collaborators and colleagues, without whom this work would not have been possible.

The work on displaced dimuons was performed in collaboration with UCLA senior researcher Slava Valuev and UCLA postdoctoral scholar Alice Florent under the supervision of my advisor, Professor Robert Cousins, along with UCLA colleagues Jay Hauser and William Nash, and Institute of High Energy Physics Vienna colleagues including Wolfgang Adam, Alberto Escalante, Suchita Kulkarni, Ivan Mikulec, Sebastian Templ, and Claudia Wulz.
We also thank our colleagues and conveners in CMS who contributed valuable knowledge and discussions, including Juliette Alimena, Paul Lujan, Ivan Mikulec, Niki Saoulidou, and Adish Vartak, and other UCLA professors Michalis Bachtis, Zvi Bern, and David Saltzberg.

The work on neutron background studies described in the appendix was performed in full collaboration with fellow UCLA graduate student Christian Schnaible under the supervision of our common advisor, Professor Robert Cousins.
Together, Christian Schnaible and I wrote a detailed draft of internal CMS Detector Note DN-2017/019, reproduced below.
Professor Cousins also provided editing assistance, particularly in \Sec~\ref{sec:datamc}.
I presented this work in a poster session at the 2017 European Physical Society Conference on High-Energy Physics (EPS-HEP), and wrote a conference note for the Proceedings.
Christian Schnaible presented this work in a talk at the the 2017 American Physical Society Meeting of the Division of Particles \& Fields (APS-DPF).

We also thank everyone on the CSC GIF++ team, the CMS UCLA group, and everyone whose work we have built on or partially reproduced and who have contributed valuable knowledge and advice, including Cameron Bravo, Tim Cox, Alice Florent, Jay Hauser, Misha Ignatenko, Vladimir Ivantchenko, Evaldas Juska, Alexey Kamenev, Andrey Korytov, \mbox{Katerina} Kuznetsova, Armando Lanaro, Paul Lujan, Alex Madorsky, Nick McColl, Hualin Mei, Yuriy Pakhotin, Vladimir Palchik, Victor Perelygin, Jian Wang, Wells Wulsin, and Piet Verwilligen.

This thesis includes material that is based upon work supported by the U.S.\ Department of Energy under Award Number {DE}-{SC}0009937.

I also thank my friends who have personally supported me over the years.
You have listened to me, provided me with understanding and caring emotional support that has kept me healthy, sane, and happy, and been the keystones of many experiences that have filled my life with vibrancy---weekends at Shawn's, Christmas parties, Friendsgivings, trivia at Mr.\ Pickwick's Pub, movie nights, S'Beer nights, Ferney markets, D\&D campaigns, French lunches, coffee and chess afternoons, and countless lunches, dinners, fireworks, BBQs, lakeside swims, wine tastings, hikes, and trips across fifteen European countries.
I cannot thank you enough for my life of the past four years; the least I can do is attempt to mention some of your names here.

\hyphenation{Capon-pon}
\hyphenation{Ban-dyo-padh-yay}
\hyphenation{Vi-van}

My high school friends including
Adam Alonzo,
Emily McGreevy Alonzo,
Danielle Caponpon,
Andy Chen,
Jeff Cruz,
Derek Dosland,
Madeline Foss,
Brandie Hanson,
Krystal Lin,
Arohi Sharma,
and
Dustin Tygret.
My Berkeley friends including
Spandan Bandyopadhyay,
Memo Carpenter,
Alexey Drobizhev,
Nitin Egbert,
Harshil Goel,
Alex Latshaw,
Kelsey Oliver-Mallory,
Michela Paganini,
and
Ana Vasudevan.
My internet friends including Katherine Rybski and Mahathi Chavali.

My UCLA friends including
Pauline Arriaga,
Cameron Bravo,
Gary Chen,
Krishna Choudhury,
Eric Cotner,
Eve Emmanouilidou,
Rhyan Ghosh,
MaryAnn Kontonicolas,
Alex Latshaw,
Nathan Majernik,
Ana Martynova,
David Perkins,
Seth Pree,
Jeff Robertson,
Chris Schnaible,
Grace Stees,
Joe Swearngin,
Shawn Tang,
and
Sam Witte.

And my CERN friends including
Juliette Alimena,
Reyer Band,
Chris Brainerd,
Cameron Bravo,
Alex Brown,
Kira Burt,
Sarah Charley,
Grace Cummings,
Andr\'{e}s Delannoy,
Pieter Everaerts,
Juliana Froggatt,
Amin Ghiasi,
Usama Hussain,
Rachel Hyneman,
Jan Kiesler,
Tony Lefeld,
Bing Liu,
Kenneth Long,
Paul Lujan,
Chris Martin,
Christine McLean,
Vivan Nguyen,
Michela Paganini,
Ashley Parker,
Michael Planer,
Brendan Regnery,
Elodie Resseguie,
Francesca Ricci-Tam,
Titas Roy,
Lesley Smith,
Nick Smith,
Eric Takasugi,
Dylan Teague,
Santona Tuli,
Ana Vasudevan,
Matthias Weber,
Nate Woods,
and
Brent Yates.

I also thank my family, without whom I would not be here. My parents, Sanjukta Dasgupta and Aloke Dasgupta, have encouraged me, supported me, instilled me with values and culture, and provided me with everything to succeed. No amount of thanks can accurately capture their impact on my life. My cousins Soham Dasgupta, Twinkle Ghosh, and Ritwika Kanungoe, who have kept me connected with a distant dream. My adopted family Sonia Das and Mihir Das, whose powerful support have formed a backbone of my life. And everyone else in my family who have showered me with love.

Last, but not least, I thank my advisor, Professor Robert Cousins.
I have learned a great deal from following Bob's guidance and listening to his advice, and his values and knowledge, often delivered as concise aphorisms, have found a permanent place in my memory.
His high standards have pushed me to learn and accomplish more than I could have expected.
And he has been generous and organized, ensuring minimal worries about logistical issues and allowing me to focus on performing the research contained within this thesis.
For all this and more I am grateful, and I consider myself lucky to have been his student.
