
\documentclass [PhD] {uclathes}

\input {commands.tex}                         % personal LaTeX macros

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Usually things live in separate flies.
%
% \input {prelim}                           % preliminary page info

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%                          PRELIMINARY PAGES                           %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title          {\large
                 Search for Long-Lived Particles Decaying to Displaced Dimuons at 13 TeV
                 and
                 Study of Neutron-Induced Background Hits\\
                 in the Muon System of the Compact Muon Solenoid}
\author         {Abhigyan Dasgupta}
\department     {Physics}
% Note:  degreeyear should be optional, but as of  5-Feb-96
% it seems required or you get a year of ``2''.   -johnh
\degreeyear     {2019}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chair          {Robert D Cousins}
\member         {Zvi Bern}
\member         {Jay Hauser}
\member         {David Saltzberg}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\dedication     {\textsl{To my parents, who lit the way}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments {\input{acknowledgments.tex}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vitaitem   {2009--2013}
            {Undergraduate Research Assistant, CONCEPT Laboratory, UC Berkeley.}
\vitaitem   {2013}
            {B.A.~(Physics) and B.A.~(Mathematics), UC Berkeley.}
\vitaitem   {2013--2015}
            {Teaching Assistant, Physics Department, UCLA.}
\vitaitem   {2014}
            {M.S.~(Physics), UCLA.}
\vitaitem   {2015--present}
            {Graduate Student Researcher, Physics Department, UCLA.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\publication    {\textsl{Title} Info}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract       {This thesis presents a search for new long-lived particles decaying to two muons
                 in the CMS detector with \pp collision data taken at $\sqrt{s} = 13\TeV$ corresponding
                 to 36.3\fbinv of integrated luminosity during Run~2 of the LHC.
                 Such decays would appear as dimuon vertices displaced with respect to the \pp interaction point.
                 The search presented in this thesis uses only the muon system in order to probe
                 the longest lifetimes to which the LHC experiments are sensitive.
                 The results are interpreted in terms of a benchmark model consisting of exotic Higgs bosons
                 decaying to long-lived scalar bosons, but are presented in an approximately model-independent
                 way suitable for reinterpretation.
                 No excess is observed.
                 The appendix describes a study of neutron-induced background in CMS endcap muon chambers.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin {document}
\phantomsection
\makeintropages

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapter_intro.tex}

\input{chapter_theory.tex}

\input{chapter_cms.tex}

\input{chapter_displaced.tex}

\input{chapter_conclusion.tex}

\appendix
\input{chapter_neutron.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{thesis}
\bibliographystyle {cmstdr_modified}

\end {document}
